Supplementary materials

### 2020-03-30 WholePlotTrajectories.tex
STOTEN article version

% User-adder preamble
\usepackage{textcomp} \usepackage{tabu} \usepackage{caption}
\captionsetup{justification = justified}
\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
\hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}

@@

\PaperTitle{Post-Disturbance Tree Community Trajectories in a Neotropical Forest} % Article title

@@

\Abstract{
In the current global change context, it is urgent to anticipate the
fate of tropical forests. This means understanding tree community
response to disturbance and the underlying processes. In that respect,
we aim here to clarify taxonomic and functional post-disturbance
trajectories, and determine the scope of the Intermediate Disturbance
Hypothesis (IDH) that remains debated in tropical forests. We analyzed
community trajectories following a disturbance gradient in a Neotropical
forest over 30 years. We considered trajectories along time of community
taxonomic and functional trajectories in terms of richness, evenness,
composition, and redundancy. We based on the annual botanical
inventories of 75 ha of a Neotropical forest and on large trait datasets
comprising seven leaf, stem, and life-history traits. We identified a
decoupling between taxonomic composition, differing among communities,
and functional composition, remaining similar and convergent. The
taxonomic diversity followed humped-shaped trajectories depending on the
disturbance intensity, which validated the IDH (Intermediate Disturbance
Hypothesis). The functional diversity trajectories,however, were
homogeneous among all plots and dismissed the IDH. We explained this
decoupling by the variations in community functional redundancy that
mitigated the functional impact of disturbance. Although consistent, the
recovery of community composition, diversity, and redundancy remained
unachieved after 30 years. These results acknowledged the need of
decades-long cycles without disturbance to ensure community complete
recovery, and questioned community resilience after repeated
disturbances.
}

@@

The large areas covered with tropical forests worldwide hold crucial
environmental, economic, and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna, and
ensure cultural and human well-being. They regulate local and regional
climates, as well as carbon, water and nutrient cycles. However, the
growing demand in forests products together with current global changes
increase the pressure on remaining undisturbed forests
\citep{Morales-Hidalgo2015}. These threats affect the natural
disturbance regime that defines and maintains the structure,
composition, and functioning of tree communities
\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. To anticipate
the fate of tropical forests, it is urgent to understand tree community
response to disturbance, and the underlying ecological processes. The
forest cover is generally maintained following disturbance, but
modifications in the fluxes of light, heat, and water
\citep{Goulamoussene2017} change community abiotic and biotic
environments. These changes translate into post-disturbance community
trajectories that have been largely studied through trajectories of
forest structural parameters such as aboveground biomass, tree height or
stem density \citep{Piponiot2016, Rutishauser2016}. Some of the
determinants of post-disturbance biomass trajectories are already
identified, like the structure and composition of the pre-disturbance
community, or the post-disturbance environmental parameters
\citep{Herault2018}. Community diversity and composition trajectories,
however, have not been as thoroughly understood
\citep{Guitet2018, Molino2001}, and manifold biodiversity trajectories
might emerge given the variety of species response to disturbance and
the diversity of tropical forests
\citep{Lindenmayer2012, Garcia_florez2017}.

An early conceptual basis of the linkage between biodiversity and
disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
states a relationship between the community diversity and the intensity
and frequency of disturbance events, and postulates a diversity peak at
intermediate level of disturbance \citep{Connell1978}. This is based on
the fluctuations of community environment following disturbance that
foster both competitively superior species and fast colonizers, and
prevents competitive exclusion \citep{Shea2004, Pulsford2016}. In
tropical forests, however, observations of the IDH often diverge from
theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}, and
the underlying processes might be complicated by the diversity of
tropical tree communities \citep{Lindenmayer2012, Garcia_florez2017}. In
this context, the IDH is controversial in tropical forests and remains
to be tested \citep{Hubbell2001, Fox2013, Sheil2013}.

@@

The experiment is a network of twelve 6.25 ha plots (Table
\ref{tab:Tab1}) that underwent three disturbance treatments in 1987
according to a randomized plot design \citep{Gourlet-Fleury2004}.\\
The experiment comprised three replicates of three sylvicultural
treatments (hereafter plots T1, T2, and T3), and three control plots
(T0). All treatments T1, T2, and T3 comprised the logging of 10 trees/ha
with 50 cm minimum DBH that belonged to a set of 58 commercially
exploited species \citep{Gourlet-Fleury2004}. Treatment T2 additionally
comprised a thinning treatment by poison-girdling of non-commercially
exploited species randomly selected with an average of 30 trees/ha with
40 cm minimum DBH. Treatment T3 additionally comprised the logging of 15
trees/ha with 40 cm minimum DBH, and the poison-girdling of 20 trees/ha
with a 50 cm minimum DBH, all belonging to non-commercially exploited
species. Considering the silvicultural treatments and the following
damage, disturbance intensity was measured as the percentage of
aboveground biomass (\%AGB) lost between the first inventory in 1984 and
five years after disturbance \citep{Piponiot2016} estimated with the
BIOMASS R package \citep{Rejou2017}. The three treatments were then
transformed into a continuous disturbance intensity gradient with
increasing above-ground biomass (AGB) loss.

@@

The study site corresponds to a tropical rainforest typical of the
Guiana Shield with a dominance of \emph{Fabaceae},
\emph{Chrysobalanaceae}, \emph{Lecythidaceae}, and \emph{Sapotaceae}. In
the 12 experimental plots, all trees above 10 cm DBH have been mapped
and measured annually since 1984. Trees are first identified with a
vernacular name assigned by the forest worker team, and afterward with a
scientific name assigned by botanists during regular botanical
campaigns. In 1984, specific vernacular names were given to 62 commmon
or commercially exploited species. More infrequent species were
identified under general identifiers only distinguishing trees and
palms. From 2003, botanical campaigns have been conducted every 5 to 6
years to identify all trees at the species level. In 2015, however,
identification levels still varied among plots and campaigns.

This variability of protocols in time raised methodological issues as
vernacular names usually correspond to various botanical species. This
resulted in significant taxonomic uncertainty that had to be accounted
for in the measure of composition and diversity metrics. Uncertainty
propagation was implented withing a Bayesian framework using
vernacular/botanical names associations to reconstitute complete
inventories at genus level from real incomplete ones. Vernacular names
were replaced through multinomial trials based on the association
probability \(\big[\alpha_1, \alpha_2,..., \alpha_V\big]\) observed
across all inventories between each vernacular name \emph{v} and all
species \(\big[s_1, s_2,..., s_N\big]\):

@@

See Supplementary Materials -Fig. S1 and \citet{Aubry-Kientz2013} for
the detailed methodology.

Six functional traits representing the leaf economics (\emph{i.e.} leaf
thickness, toughness, total chlorophyll content, and specific leaf
area), and the stem economics (\emph{i.e.} wood specific gravity and
bark thickness) were obtained from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}.
Trait values were assessed from a selection of individuals located in
nine permanent plots in French Guiana, including two in Paracou, and
comprised 294 species belonging to 157 genera. Whenever a species was in
the dataset but missed some trait values (10\% of the species), missing
values were filled using multivariate imputation by chained equation
\citep{Mice2011}. To account for the phylogenetic signal in the filling
process, imputations were based on samples of species from the same
genus or from the same family. Whenever a species was not in the
dataset, it was attributed a set of trait values randomly sampled among
species of the next higher taxonomic level (same genus or family). Two
life-history traits, maximum specific height and seed mass, came from
the Mariwenn database \footnote{https://www.ecofog.gf/mariwenn/}. The
database compiles information from a vast literature on the flora of
French Guiana \citep{Ollivier2007} and comprises 362 species belonging
to 188 genera. As seed mass information was classified into classes, no
data filling process was applied and analyses were restricted to the
botanical species recorded.

All composition and diversity metrics were obtained after 50 iterations
of the uncertainty propagation framework.

@@

Because of the variable precision of botanical identification efforts,
we had to conduct the taxonomic composition and diversity analysis at
the genus level. Taxonomic and functional trajectories of community
composition were drawn in a two-dimensional NMDS ordination plane. Two
NMDS using abundance-based (\emph{i.e.} Bray-Curtis) dissimilarity
measures were conducted to map either taxonomic or functional
composition, the latter based on the seven leaf, stem, and life history
traits (without seed mass classes). Trajectories along time were
reported through the Euclidean distance between the target inventories
and the 1984 pre-disturbance inventories of reference. Univariate
trajectories of the leaf, stem, and life-history traits were also
visualized with the community weighted means (CWM) \citep{Diaz2007}.
Species seed mass were given in 5 mass classes, and seed mass
trajectories were reported as the proportion of each class in the
inventories (Supplementary materials).

The taxonomic and functional trajectories were analysed from the 1984
pre-disturbance inventories of reference. The taxonomic diversity was
reported through species richness and the Hill number translation of the
Simpson index \citep{Hill1973}. The comparison between these two metrics
assesses community taxonomic richness and evenness: thereafter, results
will be discussed directly in terms of taxonomic richness and evenness.
Both indices are recommended for diversity studies \citep{Marcon2015},
and belong to the set of HCDT or generalized entropy corresponding,
respectively, to the 0 and 2 order of diversity (q). The functional
diversity was reported using the functional richness and functional
evenness, through the Rao index of quadratic entropy. The Rao index
combines species abundance distribution, and the average pairwise
functional dissimilarity between species computed by the Gower distance.

The impacts of the initial disturbance levels on the maximum gain or
loss in richness and evenness were tested with the Spearman rank
correlation tests. Richness and evenness trajectories were besides
analyzed through polynomial regression between (i) taxonomic and
functional richness and evenness, and (ii) the initial \%AGB loss at 10,
20, and 30 years after disturbance.

Finally, functional redundancy was measured as the overlap among species
in community functional space \citep{Carmona2016}. First, the
individuals of the trait database were mapped in the plane of the first
two axes from a PCA analysis. The PCA analysis lowered the weight of
correlations among traits as the axes are combinations of the most
decoupled traits. For each species, the traits probability density (TPD)
were computed from the mapping of individuals through two-dimension
kernel density estimators. Second, for each community, the TPD weigthed
by species abundance were summed accross the functional space. Third,
the functional space was divided into a 100 x 100 grid, and the number
of species with a positive TPD was counted in each cell. The average
count across cells minus 1 returned the Community Functional Redundancy,
which was the average number of species in the community that share the
same trait values.

@@

The functional composition trajectory resembled, in disturbed
communities, to cyclic compositional changes with an incomplete recovery
of the initial composition (Fig.\ref{fig:NMDSplans}). The maximum
dissimilarity with the initial state was positively correlated with the
disturbance intensity for both taxonomic and functional composition
(\(\rho_{Spearman}^{Taxonomic}=0.87\) and
\(\rho_{Spearman}^{Functional}=0.90\), respectively). The maximum
dissimilarity with the initial state was reached for taxonomic
composition between 15 to 25 years, and between 20 to 25 years for
functional composition.

@@

In control plots, Community Weighted Means (CWM) of functional traits
remained stable in time.\\
In disturbed plots, they mostly followed unimodal trajectories, either
stabilizing or returning towards their initial values, to the exception
of leaf chlorophyll content, which continued to increase 30 years after
disturbance for 4 out of 6 highly disturbed plots. Maximum height at
adult stage (\emph{Hmax}), leaf toughness, and wood specific gravity
(\emph{WSG}) decreased in time and then slightly increased, but remained
significantly lower than their initial value (Fig. \ref{fig:CWM}). Bark
thickness and specific leaf area (\emph{SLA}) both increased along time.
Bark thickness remained substantially high after 30 years, and
\emph{SLA} had almost recovered to its initial value. Whatever the
functional traits, the maximum difference to initial value was highly
correlated to the disturbance intensity. Positive correlations were
observed for leaf thickness, chlorophyll content, SLA and bark thickness
(\(\rho_{Spearman}^{Leaf thickness}=0.76\),
\(\rho_{Spearman}^{Chlorophyll content}=0.60\),
\(\rho_{Spearman}^{SLA}=0.93\),
\(\rho_{Spearman}^{Bark thickness}=0.71\)). Negative correlation was
observed for leaf toughness, WSG, and Hmax
(\(\rho_{Spearman}^{Leaf toughness}=-0.53\),
\(\rho_{Spearman}^{WSG}=-0.75\), \(\rho_{Spearman}^{Hmax}=-0.40\)) The
proportions of the three lightest seed mass classes increased in all
disturbed plots. After 30 years the proportion of lightest seed mass
class decreased while it stabilized for the two other lightest seed mass
classes (Supp. Mat. - Fig. S2).

@@

Taxonomic richness and evenness remained stable in control plots over
the 30 years of monitoring. In disturbed communities, after low
disturbance intensity the taxonomic richness increased, reaching a
maximum gain of 14 botanical genera. After intense disturbance the
taxonomic richness followed a more complex trajectory, decreasing for
ten years after disturbance before recovering to pre-disturbance values.
The maximum richness loss or gain after disturbance was positively
correlated with the disturbance intensity
(\(\rho_{Spearman}^{Richness}=0.50\)).

In all disturbed plots the evenness first increased until a maximum
reached after around 20 years. This maximum was positively correlated
with the disturbance intensity (\(\rho_{Spearman}^{Simpson}=0.77\)). The
evenness then stabilized except for two intensively-disturbed plots
(number 8 and 12) for which it kept increasing (Fig. \ref{fig:DivTaxo}).

@@

\caption{Trajectories of community taxonomic richness \textbf{(a)}, Simpson diversity \textbf{(b)}, functional richness \textbf{(c)}, and Rao diversity \textbf{(d)}. Values correspond to the difference over 30 years of community diversity with the values of 1984 pre-disturbance inventories of reference. Shaded areas are the credibility intervals }\label{fig:DivTaxo}
\end{figure*}

Functional richness and evenness remained stable in control plots over
the 30 years of monitoring. In disturbed communities, both trajectories
depended on the disturbance intensity, with their maximum values in time
being positively correlated to \%AGB loss
\(\rho_{Spearman}^{Richness} = 0.76\) and\\
\(\rho_{Spearman}^{Rao} = 0.60\). For low disturbance intensity,
functional richness and evenness displayed a low but long-lasting
increase up to a maximum reached after 20-25 years. For high disturbance
intensity, they generally displayed a fast but short increase followed,
after 10 years, by a slow decrease towards the initial values.

The second-degree polynomial regressions between (i) the percentage AGB
loss, and (ii) the taxonomic and functional diversity showed various
shapes depending on the diversity indices and on the time since
disturbance (Fig. \ref{fig:IDHplot}). Regarding taxonomic diversity, the
relationship between disturbance intensity and diversity was more
markedly hump-shaped for richness than for evenness, and peaked at 20\%
of initial AGB loss. Regarding functional diversity, the relationship
was almost linear, and was similar between richness and evenness. All
the relationships between disturbance intensity and diversity were
stronger 20 or 30 years after disturbance than they were 10 years after
disturbance.

@@

Control plots displayed stable functional redundancy over time. All
disturbed plots had lower functional redundancy than control plots and
followed similar hump-shaped trajectories (Fig. \ref{fig:RedFunRest}).
The maximum redundancy loss was positively correlated with the
disturbance intensity (\(\rho_{Spearman}=0.47\)) and the recovery
trajectory had not attained initial values for any disturbed communities
after 30 years.

@@

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/RedFunRest-1}

@@

\caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance. Shaded areas are the credibility intervals.}\label{fig:RedFunRest}
\end{figure}

>>>>

% User-adder preamble
\usepackage{textcomp} \usepackage{tabu} \usepackage{caption}
\captionsetup{justification = justified}
\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
\hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
\usepackage[left]{lineno} \linenumbers \usepackage{bookmark}

@@

\PaperTitle{Diverging taxonomic and functional trajectories following disturbance in
a Neotropical forest} % Article title

@@

\Abstract{
In the current global change context, it is urgent to anticipate the
fate of tropical forests. This means understanding tree community
response to disturbance and the underlying processes. In that respect,
we aim here to clarify taxonomic and functional post-disturbance
trajectories, and determine the scope of the Intermediate Disturbance
Hypothesis (IDH) that remains debated in tropical forests.
\color{black}We analyzed community trajectories following a disturbance
gradient from 10 to 60\% of above-ground biomass loss in a Neotropical
forest over 30 years.\color{red} We considered trajectories along time
of community taxonomic and functional trajectories in terms of richness,
evenness, composition, and redundancy. We based on the annual botanical
inventories of 75 ha of a Neotropical forest and on large trait datasets
comprising seven leaf, stem, and life-history traits. We identified a
decoupling between taxonomic composition, differing among communities,
and functional composition, \color{red} similar among communities and
convergent in the functional space.\color{black} \color{red} The
taxonomic diversity followed humped-shaped trajectories along time after
disturbance depending on the initial disturbance intensity, which
validated the IDH (Intermediate Disturbance Hypothesis).\color{black}
The functional diversity trajectories,however, were homogeneous among
plots and dismissed the IDH. We explained this decoupling by the
variations in community functional redundancy that mitigated the
functional impact of disturbance. Although consistent, the recovery of
community composition, diversity, and redundancy remained unachieved
after 30 years. \color{red} These results acknowledged the need of
decades-long cycles without disturbance to ensure community complete
recovery.\color{black}
}

@@

The large areas covered with tropical forests worldwide hold crucial
environmental, economic, and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna, and
ensure cultural and human well-being. They regulate local and regional
climates, as well as carbon, water and nutrient cycles. However, the
growing demand in forests products together with current global changes
increase the pressure on remaining undisturbed forests
\citep{Morales-Hidalgo2015}. \color{red}These threats may change the
frequency and magnitude of the natural disturbance regime that defines
and maintains the structure, composition, and functioning of tree
communities \citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}.
\color{black} To anticipate the fate of tropical forests, it is urgent
to understand tree community response to disturbance, and the underlying
ecological processes. The forest cover is generally maintained following
disturbance, but modifications in the fluxes of light, heat, and water
\citep{Goulamoussene2017} change community abiotic and biotic
environments. These changes translate into post-disturbance community
trajectories that have been largely studied through trajectories of
forest structural parameters such as aboveground biomass, tree height or
stem density \citep{Piponiot2016, Rutishauser2016}. Some of the
determinants of post-disturbance biomass trajectories are already
identified, like the structure and composition of the pre-disturbance
community, or the post-disturbance environmental parameters
\citep{Herault2018}. Community diversity and composition trajectories,
however, have not been as thoroughly understood
\citep{Guitet2018, Molino2001}, and manifold biodiversity trajectories
might emerge given the variety of species response to disturbance and
the diversity of tropical forests
\citep{Lindenmayer2012, Garcia_florez2017}.

An early conceptual basis of the linkage between biodiversity and
disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
states a relationship between the community diversity and the intensity
and frequency of disturbance events, and postulates a diversity peak at
intermediate level of disturbance \citep{Connell1978}. This is based on
the fluctuations of community environment following disturbance that
foster both competitively superior species and fast colonizers, and
prevents competitive exclusion \citep{Shea2004, Pulsford2016}.
\color{red} In tropical forests, some studies advocate community
deterministic response to disturbance, as for example in the dry and wet
forests in Ghana \citep{Bongers2009} or in the guianese rainforests
\citep{Molino2001}, where intermediate disturbance proved to increase
the community richness by increasing the proportion of pioneers or
heliophilous species in the community while maintaining the richness in
old-growth forest species. In other cases, however, observations of the
IDH diverge from theoretical expectations
\citep{Hugues2007, Sheil2003, Norden2017}. Some studies refute the role
of disturbance intensity, as observed in forest gaps in Panama where no
difference in species richness was found between forest gaps and
undisturbed areas \citep{Hubbell2001}, or in costa-rican forests where
forests that did not follow deterministic convergence display stochastic
trajectories \citep{Norden2015}.\\
The high diversity of tropical forests may foster the emergence of
numerous facilitation, adaptation, and inter- and intra-specific
competition following disturbance
\citep{Garcia_florez2017, Bongers2009}.These interactions can result in
miscellaneous responses to disturbance \citep{Lindenmayer2012}, which
question the validity of the IDH in tropical forests
\citep{Hubbell2001, Fox2013, Sheil2013}. \color{black}

@@

The experiment is a network of twelve 6.25 ha plots\color{red},
distributed over a 400ha area \color{black}(Table \ref{tab:Tab1}, Supp.
Mat. Fig. S1), that undergone three disturbance treatments in 1987
according to a randomized plot design \citep{Gourlet-Fleury2004}.\\
\color{red}The experiment intended to simulate logging intensities and
distinguished the treatments for non-commercial commercial species,
logged at 40cm DBH to simulate thinning operations.\color{black} The
experiment comprised three replicates of three sylvicultural treatments
(hereafter plots T1, T2, and T3), and three control plots (T0).
\color{red}Control plots are set in mature forest stand that did not
undergone any logging nor intense anthropic activities activities for
the last 500 years\color{black}. All treatments T1, T2, and T3 comprised
the logging of 10 trees/ha with 50 cm minimum DBH that belonged to a set
of 58 commercially exploited species \citep{Gourlet-Fleury2004}.
Treatment T2 additionally comprised a thinning treatment by
poison-girdling of non-commercially exploited species randomly selected
with an average of 30 trees/ha with 40 cm minimum DBH. Treatment T3
additionally comprised the logging of 15 trees/ha with 40 cm minimum
DBH, and the poison-girdling of 20 trees/ha with a 50 cm minimum DBH,
all belonging to non-commercially exploited species. Considering the
silvicultural treatments and the following damage, disturbance intensity
was measured as the percentage of aboveground biomass (\%AGB) lost
between the first inventory in 1984 and five years after disturbance
\citep{Piponiot2016} estimated with the BIOMASS R package
\citep{Rejou2017}. \color{red} Tree diameter are provided in the
experimental dataset, wood density is extracted from global wood density
(GWD) database \citep{Zanne2009}, and tree height is retrieved from a
generic H--D model based on a single bioclimatic predictor (eqn 6a in
\citep{Chave2014}). The AGB is measured from the generalized allometric
model eqn 4 of \citep{Chave2014}. \color{black} The three treatments
were then transformed into a continuous disturbance intensity gradient
with increasing above-ground biomass (AGB) loss.

@@

The study site corresponds to a tropical rainforest typical of the
Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae,
Lecythidaceae, and Sapotaceae. In the 12 experimental plots, all trees
above 10 cm DBH have been mapped and measured annually since 1984. Trees
are first identified with a vernacular name assigned by the forest
worker team, and afterward with a scientific name assigned by botanists
during regular botanical campaigns. In 1984, specific vernacular names
were given to 62 commmon or commercially exploited species. More
infrequent species were identified under general identifiers only
distinguishing trees and palms. From 2003, botanical campaigns have been
conducted every 5 to 6 years to identify all trees at the species level.
In 2015, however, identification levels still varied among plots and
campaigns.

This variability of protocols in time raised methodological issues as
vernacular names usually correspond to various botanical species. This
resulted in significant taxonomic uncertainty that had to be accounted
for in the measure of composition and diversity metrics. Uncertainty
propagation was implented within a Bayesian framework using
vernacular/botanical names associations to reconstitute complete
inventories at genus level from real incomplete ones. Vernacular names
were replaced through multinomial trials based on the association
probability \(\big[\alpha_1, \alpha_2,..., \alpha_V\big]\) observed
across all inventories between each vernacular name \emph{v} and all
species \(\big[s_1, s_2,..., s_N\big]\):

@@

See Supplementary Materials -Fig. S2 and \citet{Aubry-Kientz2013} for
the detailed methodology.

Six functional traits representing the leaf economics (\emph{i.e.} leaf
thickness, toughness, total chlorophyll content, and specific leaf
area), and the stem economics (\emph{i.e.} wood specific gravity and
bark thickness) were obtained from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}.
Trait values were assessed from a selection of individuals located in
nine permanent plots in French Guiana, including two in Paracou, and
comprised 294 species belonging to 157 genera. Whenever a species was in
the dataset but missed some trait values (10\% of the species), missing
values were filled using multivariate imputation by chained equation
\citep{Mice2011}. To account for the phylogenetic signal in the
imputation process, imputations were based on samples of species from
the same genus or from the same family. Whenever a species was not in
the dataset, it was attributed a set of trait values randomly sampled
among species of the next higher taxonomic level (same genus or family).
Two life-history traits, maximum specific height and seed mass, came
from the Mariwenn database \footnote{https://www.ecofog.gf/mariwenn/}.
The database compiles information from a vast literature on the flora of
French Guiana \citep{Ollivier2007} and comprises 362 species belonging
to 188 genera. As seed mass information was classified into classes, no
data filling process was applied and analyses were restricted to the
botanical species recorded.

\color{red}Confidence intervals of the composition and diversity metrics
were computed after 50 iterations of the uncertainty propagation
framework.\color{black}

@@

Because of the variable precision of botanical identification efforts,
we had to conduct the taxonomic composition and diversity analysis at
the genus level. \color{red}Functional composition, diversity, and
redundancy measures were conducted at species level.\color{black}
Taxonomic and functional trajectories of community composition were
drawn in a two-dimensional NMDS ordination plane. Two NMDS using
abundance-based (\emph{i.e.} Bray-Curtis) dissimilarity measures were
conducted to map either taxonomic or functional composition, the latter
based on the seven leaf, stem, and life history traits (without seed
mass classes). Trajectories along time were reported through the
Euclidean distance between the target inventories and the 1984
pre-disturbance inventories of reference. Univariate trajectories of the
leaf, stem, and life-history traits were also visualized with the
community weighted means (CWM) \citep{Diaz2007}. Species seed mass were
given in 5 mass classes, and seed mass trajectories were reported as the
proportion of each class in the inventories (Supplementary materials).

The taxonomic and functional trajectories were analysed from the 1984
pre-disturbance inventories of reference. The taxonomic diversity was
reported through species richness and the Hill number translation of the
Simpson index \citep{Hill1973}. The comparison between these two metrics
assesses community taxonomic richness and evenness: thereafter, results
will be discussed directly in terms of taxonomic richness and evenness.
Both indices are recommended for diversity studies \citep{Marcon2015},
and belong to the set of \color{red} Tsallis or generalized entropy
\color{black} corresponding, respectively, to the 0 and 2 order of
diversity (q). \color{red}Diversity metrics were measured using the Hqz
function from the entropart package \citep{Marcon2015}.\color{black} The
functional diversity was reported using the functional richness and
functional evenness. \color{red}Both functional richness and evenness
were measured through a distance-based dissimilarity matrix built with
the daisy function of the cluster R package \citep{Cluster2019}, using
the Gower distance between species, similar to \citet{Herault2007a},
\citet{Herault2007b}.\color{black} The functional diversity was measured
through the Rao index of quadratic entropy. The Rao index combines the
community abundance distribution\color{red}, from species relative
abundance, \color{black} and the average pairwise functional
dissimilarity between species computed by the Gower distance.
\color{red}The comparison between these two metrics assesses community
functional richness and evenness: thereafter, results will be discussed
directly in terms of functional richness and evenness.\color{black}

The impacts of the initial disturbance levels on the maximum gain or
loss in richness and evenness were tested with the Spearman rank
correlation tests. Richness and evenness trajectories were besides
analyzed through polynomial regression \color{red}or simple linear
models \color{black} between (i) taxonomic and functional richness and
evenness, and (ii) the initial \%AGB loss at 10, 20, and 30 years after
disturbance. \color{red}The best model fitting, between linear or
quadratic model, was chosen based on their respective AIC criterion. The
explanatory power of the fit were measured to adjusted r-square
values.\color{black}

Finally, functional redundancy was measured as the overlap among species
in community functional space \citep{Carmona2016}. First, the
individuals of the trait database were mapped in the plane of the first
two axes from a PCA analysis. The PCA analysis lowered the weight of
correlations among traits as the axes are combinations of the most
decoupled traits. \color{red}The two main PCA axis respectively hold
26\% and 19\% of the trait dataset variance.\color{black} For each
species, the traits probability density (TPD) were computed from the
mapping of individuals through two-dimension kernel density estimators.
\color{red} We used the kde2d function of the MASS package
\citep{MASS2002} that is a gaussian density estimator for a
2-dimensional space. \color{black} Second, for each community, the TPD
weigthed by species abundance were summed across the functional space.
Third, the functional space was divided into a 100 x 100 grid, and the
number of species with a positive TPD was counted in each cell. The
average count across cells minus 1 returned the Community Functional
Redundancy, which was the average number of species in the community
that share the same trait values.

@@

The functional composition trajectory resembled, in disturbed
communities, to cyclic compositional changes with an incomplete recovery
of the initial composition (Fig.\ref{fig:NMDSplans}). The maximum
dissimilarity with the initial state was positively correlated with the
disturbance intensity for both taxonomic and functional composition
(\(\rho_{Spearman}^{Taxonomic}=0.94\) and
\(\rho_{Spearman}^{Functional}=0.90\), respectively. Fig.
\ref{fig:SumUp}). The maximum dissimilarity with the initial state was
reached for taxonomic composition between 15 to 25 years, and between 20
to 25 years for functional composition.

@@

In control plots, Community Weighted Means (CWM) of functional traits
remained stable in time.\\
In disturbed plots, they mostly followed unimodal trajectories, either
stabilizing or returning towards their initial values, to the exception
of leaf chlorophyll content, which continued to increase 30 years after
disturbance for 4 out of 6 highly disturbed plots. \color{red}Main
tendencies for disturbed plots were a decrease followed by a slight
increase of maximum height at adult stage (\emph{Hmax}), leaf toughness,
and wood specific gravity (\emph{WSG}), but trait values remained
significantly lower than their initial value (Fig. \ref{fig:CWM}).
\color{black} Bark thickness and specific leaf area (\emph{SLA}) both
increased along time. Bark thickness remained substantially high after
30 years, and \emph{SLA} had almost recovered to its initial value.
Whatever the functional traits, the maximum difference to initial value
was highly correlated to the disturbance intensity. \color{red}Positive
correlations were observed for leaf thickness, chlorophyll content, SLA
and bark thickness (\(\rho_{Spearman}^{Leaf thickness}=0.79\),
\(\rho_{Spearman}^{Chlorophyll content}=0.61\),
\(\rho_{Spearman}^{SLA}=0.99\),
\(\rho_{Spearman}^{Bark thickness}=0.68\)). Maxima were observed between
5 and 15 years following disturbance for leaf thickness and SLA while
chlorophyll content and bark thickness kept increasing after 30 years
for leaf chlorophyll content and bark thickness. Negative correlation
was observed for leaf toughness, WSG, and Hmax
(\(\rho_{Spearman}^{Leaf toughness}=-0.47\),
\(\rho_{Spearman}^{WSG}=-0.80\), \(\rho_{Spearman}^{Hmax}=-0.28\)).
Minima were observed between 5 and 15 years following disturbance for
Hmax, 10 and 20 years for WSG, and between 15 and 25 years following
disturbance for leaf toughness.\color{black} The proportions of the
three lightest seed mass classes increased in all disturbed plots. After
30 years the proportion of lightest seed mass class decreased while it
stabilized for the two other lightest seed mass classes (Supp. Mat. -
Fig. S3).

@@

Taxonomic richness and evenness remained stable in control plots over
the 30 years of monitoring. In disturbed communities, after low
disturbance intensity the taxonomic richness increased, reaching a
maximum gain of 14 botanical genera. After intense disturbance the
taxonomic richness followed a more complex trajectory, decreasing for
ten years after disturbance before recovering to pre-disturbance values.
\color{red}The maximum rate of change compared to initial richness was
positively correlated with the disturbance intensity
(\(\rho_{Spearman}^{Richness}=0.73\)).\color{black}

In all disturbed plots \color{red}, to the exception of plots 5 (T2) and
12 (T3) that showed a small initial drop, \color{black} the evenness
first increased until a maximum reached after around 20 years. This
maximum was positively correlated with the disturbance intensity
(\(\rho_{Spearman}^{Evenness}=0.82\), \color{red} Fig.
\ref{fig:SumUp})\color{black}. The evenness then stabilized except for
two intensively-disturbed plots (number 8 and 12) for which it kept
increasing (Fig. \ref{fig:DivTaxo}).

@@

\caption{Trajectories of community taxonomic richness \textbf{(a)}, taxonomic evenness, \textbf{(b)}, functional richness \textbf{(c)}, and functional evenness \textbf{(d)}. Values correspond to the difference over 30 years of community diversity with the values of 1989 inventories of reference, 5 years after disturbance. Shaded areas are the credibility intervals }\label{fig:DivTaxo}
\end{figure*}

Functional richness and evenness remained stable in control plots over
the 30 years of monitoring. \color{red}In disturbed communities, both
functional richness and evenness trajectories increased following
disturbance, with maxima positively correlated to \%AGB loss
\(\rho_{Spearman}^{Richness} = 0.76\) and
\(\rho_{Spearman}^{Evenness} = 0.60\) (Fig. \ref{fig:SumUp}).
\color{black} For low \color{red}and intermediate
\color{black}disturbance intensity, functional richness and evenness
displayed a low but long-lasting increase up to a maximum reached after
20-25 years. For high disturbance intensity, they generally displayed a
fast but short increase followed, after 10 years, by a slow decrease
towards the initial values.

The second-degree polynomial regressions between (i) the percentage AGB
loss, and (ii) the taxonomic and functional diversity showed various
shapes depending on the diversity indices and on the time since
disturbance (Fig. \ref{fig:IDHplot}). \color{red}Regarding taxonomic
diversity, the taxonomic richness showed a humped-shaped trajectory with
the disturbance intensity, and peaked at 20\% of initial AGB loss. The
relationship between disturbance intensity and taxonomic evenness,
however, was humped-shaped only 20 years after disturbance and linear
for 10 and 30 years after disturbance times.\color{black} Regarding
functional diversity, \color{red} the relationship was linear for both
taxonomic richness and evenness, except for the functional diversity
after 30 years. \color{black}

@@

Control plots displayed stable functional redundancy over time. All
disturbed plots had lower functional redundancy than control plots and
followed similar humped-shaped trajectories (Fig. \ref{fig:RedFunRest}).
The maximum redundancy loss was positively correlated with the
disturbance intensity (\(\rho_{Spearman}=0.31\), (Fig. \ref{fig:SumUp}))
and the recovery trajectory had not attained initial values for any
disturbed communities after 30 years.

@@

{\centering \includegraphics[width=0.4\linewidth]{WholePlotTrajectories_files/figure-latex/RedFunRest-1}

@@

\caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance. The redundancy is measured in a two-dimensional plan summarizing 45\% of species variance in the functional space of the 7 leaf, stem, and life history traits. Redundancy value represent the average number of species in the community that share the same trait values. Shaded areas are the credibility intervals.}\label{fig:RedFunRest}
\end{figure}

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/SumUp-1}

\caption{Maximum changes in taxonomic and functional diversity and composition against disturbance intensity, expressed as percentage of initial disturbance. Upper panels display taxonomic richness and evenness \textbf{(a)}, and functional richness and evenness \textbf{(b)} expressed in percentage of change compared to initial value. Lower panels display taxonomic and functional euclidean distance from pre-disturbance inventory, respectively in the spaces of species and functional traits \textbf{(c)}, and functional redundancy expressed as number of redundant species in the community \textbf{(d)}.}\label{fig:SumUp}
\end{figure}
